---
title: "Exploratory analysis"
author: "Nathalie Fadel"
date: "11/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read in main data set from fivethirtyeight github repository on police killings.
```{r}
library(tidyverse)
require(RCurl)

raw_data = read_csv(
        "https://raw.githubusercontent.com/fivethirtyeight/data/master/police-killings/police_killings.csv"
      )
```

A look into the cities...
```{r}
city_data = raw_data %>% 
  mutate(city_state = paste(city, state, sep = ", ")) %>% 
  group_by(city_state) %>% 
  summarise(total_killed = n())
```

Found that most cities have 1 killing in 2015. LA had the most with 9. This is not promising for an analysis that has depth of understanding. 

Looking to download data from same resource for 2016. The website has a downloadable csv. Formatted exactly the same. Only missing census data. 

```{r}
raw_2016 = read_csv("./data/the-counted-2016.csv")

```

Just looking at the raw data there is a huge increase from 467 killings in 2015 to 1093 in 2016. 

```{r}
city_data_16 = raw_2016 %>% 
  mutate(city_state = paste(city, state, sep = ", ")) %>% 
  group_by(city_state) %>% 
  summarise(total_killed = n())
```

LA stays the top but more than doubles in deaths. Houston and Pheonix trade places but remain 2 and 3. All the numbers for the top have double or quadroupled. 