---
title: "Exploratory analysis"
author: "Jaclyn Verity"
date: "11/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#First look at City level 

Read in main data set from fivethirtyeight github repository on police killings.
```{r github data}
library(tidyverse)
require(RCurl)

raw_data = read_csv(
        "https://raw.githubusercontent.com/fivethirtyeight/data/master/police-killings/police_killings.csv"
      )
```

A look into the cities...
```{r clean github}
city_data = raw_data %>% 
  mutate(city_state = paste(city, state, sep = ", ")) %>% 
  group_by(city_state) %>% 
  summarise(total_killed = n())
```

Found that most cities have 1 killing in 2015. LA had the most with 9. This is not promising for an analysis that has depth of understanding. 

Looking to download data from same resource for 2016. The website has a downloadable csv. Formatted exactly the same. Only missing census data. 

```{r load 2016 data}
raw_2016 = read_csv("./data/the-counted-2016.csv")

```

Just looking at the raw data there is a huge increase from 467 killings in 2015 to 1093 in 2016. 

```{r clean 2016 data}
city_data_16 = raw_2016 %>% 
  mutate(city_state = paste(city, state, sep = ", ")) %>% 
  group_by(city_state) %>% 
  summarise(total_killed = n())
```

LA stays the top but more than doubles in deaths. Houston and Pheonix trade places but remain 2 and 3. All the numbers for the top have double or quadroupled. 

```{r combine data}
raw_2016 = raw_2016 %>% 
  mutate(cause = classification)

raw_combined = bind_rows(raw_data, raw_2016) %>% 
  mutate(month = factor(month, levels = c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")),
         year = as.integer(year))

city_data = city_data %>% 
  mutate(year = 2015)

city_data_16 = city_data_16 %>% 
  mutate(year = 2016)

city_combined = bind_rows(city_data, city_data_16) %>% 
  mutate(year = as.factor(year))
```

```{r plot killings}
top_cities = city_combined %>% 
 top_n(20, total_killed)

ggplot(top_cities, aes(fill = year, x = city_state, y = total_killed)) + geom_bar(position = "dodge", stat = "identity") +
  theme(axis.text.x = element_text(angle = 80, hjust = 1))
```

Most of the Top 20 for police killings occured in 2016. The top 3 in 2016 also made the Top 20 overall. Looks like # 1 in 2015 would be 6 in 2016. 

#Cities in Depth
```{r}
city_depth_data = raw_combined %>% 
  filter(city %in% c( "Los Angeles", "Houston", "Phoenix", "Chicago", "San Antonio")) %>% 
  mutate(cause = as.factor(cause),
         armed = as.factor(armed))

plot_cause = city_depth_data %>% 
  select(city, cause, year) %>% 
  group_by(city, cause) %>% 
  summarise(freq = n()) %>% 
  spread(key = cause, value = freq) 

plot_cause[is.na(plot_cause)] <- 0

plot_cause1 = plot_cause %>% 
  janitor::clean_names() %>% 
  gather(key = cause, value = freq, death_in_custody:taser) %>% 
  ggplot(aes(fill = cause, x = city, y = freq)) + geom_bar(position = "dodge", stat = "identity") + labs(x = "City", y = "Frequency", title = "Cause of Death by Police Weapon")
  
  


```

