---
title: "Time Dependent Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

```{r, echo = FALSE, message = FALSE}

raw_data = read_csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/police-killings/police_killings.csv") %>%
  janitor::clean_names()

clean_data = raw_data %>%
  mutate(month = factor(month, levels = c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")),
         year = as.integer(year))

date_data = clean_data %>%
  mutate(month = as.integer(month)) %>%
  mutate(date = paste(month, day, year, sep = "-")) %>%
  mutate(date = as.Date(date, "%m-%d-%Y"))
```
We read in the csv file and made month a leveled factor variable. This data was read into figure 1 to be grouped by month. Furthermore, this dataset was then changed to make month into an integer, a new date variable was created, and date was formatted as numerical month-day-year. This data was then used to make figure 2. 

#Plots
```{r, echo = FALSE}
clean_data %>%
  group_by(month) %>%
  summarise(total_killed = n()) %>%
  ggplot(aes(x = month, y = total_killed)) +
  geom_col(size = 3) +
  labs(x = "Month", y = "number", 
       caption = "Fig. 1: Number of police killings by month, Jan-June 2015") 

date_data %>%
  group_by(date) %>%
  summarise(total_killed = n()) %>%
  ggplot(aes(x = date, y = total_killed)) +
  geom_line(size = 0.5) +
  labs(x = "date", y = "number", 
       caption = "Fig. 2: Number of police killings over time, Jan-June 2015")

```
  
As we can see from figures 1 and 2, March 2015 had the highest number of killings out of this 6 month time period, peaking late in the month. 