---
title: "The Top 5 Cities"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}

library(tidyverse)
library(reshape2)
require(RCurl)

raw_data = read_csv(
        "https://raw.githubusercontent.com/fivethirtyeight/data/master/police-killings/police_killings.csv"
      )

clean_2015 = raw_data %>% 
  mutate(month = factor(month, levels = c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")),
         year = as.integer(year))

city_depth_data = clean_2015 %>% 
  filter(city %in% c( "Los Angeles", "Houston", "Phoenix", "Chicago", "San Antonio")) %>% 
  mutate(cause = as.factor(cause),
         armed = as.factor(armed))
```


While we hope to identify patterns across the US, the sample in most cities is relatively small. Here we take an interesting look into the 5 cities with the most police killings in 2015. 

```{r  echo=FALSE}
top_5_cities = city_depth_data %>% 
  group_by(city) %>% 
  summarise(total_killed = n()) %>% 
  top_n(5, total_killed) %>% 
  arrange(-total_killed) %>%
  rename(City = city) %>% 
  rename(Killed = total_killed) %>% 
  knitr::kable()

top_5_cities
```

Most of us assume police kill with their guns. This graph shows us that the assumption is mostly accurate. In the top 5 cities for police killings death by gunshot was by far the most common cause of death for the victim. 

```{r echo=FALSE}
plot_cause = city_depth_data %>% 
  select(city, cause, year) %>% 
  group_by(city, cause) %>% 
  summarise(freq = n()) %>% 
  spread(key = cause, value = freq) 

plot_cause[is.na(plot_cause)] <- 0

plot_cause1 = plot_cause %>% 
  janitor::clean_names() %>% 
  gather(key = cause, value = freq, gunshot:taser) %>% 
  ggplot(aes(fill = cause, x = city, y = freq)) + geom_bar(position = "dodge", stat = "identity") + labs(x = "City", y = "Frequency", title = "Cause of Death by Police Weapon", caption = "Jan-Jun 2015") + scale_fill_brewer(palette = "BrBG") + 
  theme(panel.background = element_rect(fill = "lightblue",
                                colour = "lightblue",
                                size = 0.5, linetype = "solid"))
  
plot_cause1
```

Another point of tension portrayed in the media is whether or not the victim was armed. Here we can see that the victim was often unarmed. In Houston all but 1 victim were unarmed. In San Antonio the victim was never unarmed. 

```{r echo=FALSE}
plot_armed = city_depth_data %>% 
  select(city, armed) %>% 
  group_by(city, armed) %>% 
  summarise(freq = n()) %>% 
  spread(key = armed, value = freq) 

plot_armed[is.na(plot_armed)] <- 0

plot_armed1 = plot_armed %>% 
  janitor::clean_names() %>%
  rename(unarmed = no) %>% 
  gather(key = armed, value = freq, firearm:vehicle) %>% 
  ggplot(aes(fill = armed, x = city, y = freq)) + geom_bar(position = "dodge", stat = "identity") + labs(x = "City", y = "Frequency", title = "Victim Armed with a Weapon", caption = "Jan - Jun 2015") + scale_fill_brewer(palette = "BrBG") +
  theme(panel.background = element_rect(fill = "lightblue",
                                colour = "lightblue",
                                size = 0.5, linetype = "solid"))
  
plot_armed1
```

